Hydra Battles and Cie

### 2021-08-10 doc/part-hydras.tex
initialize the branch Bridge

\index{Coq!Unicity of equality proofs}
\label{sect:eq-proof-unicity}

@@

\begin{Coqsrc}
Class E0 : Type := t1_2o {cnf : T1; cnf_ok : nf cnf}.
\end{Coqsrc}

@@

\begin{Coqsrc}
Declare Scope E0_scope.
Delimit Scope E0_scope with e0.
Open Scope E0_scope.
\end{Coqsrc}

@@

\begin{Coqsrc}
Definition Lt (alpha beta : E0) := T1.LT (@cnf alpha) (@cnf beta).
Definition Le (alpha beta : E0) := T1.LE (@cnf alpha) (@cnf beta).

Infix "o<" := Lt : E0_scope.
Infix "o<=" := Le : E0_scope.
\end{Coqsrc}

@@

\begin{Coqsrc}
Require Import Logic.Eqdep_dec.

Lemma nf_proof_unicity :
  forall (alpha:T1) (H H': nf alpha), H = H'.

Lemma E0_eq_iff alpha beta : alpha = beta <-> cnf alpha = cnf beta.
\end{Coqsrc}

@@

\label{sect:omega-T1}
%\index{Constants!zero:T1}

\begin{Coqsrc}
Instance Zero : E0.
Proof.
  now exists T1.zero.
Defined.

Instance _Omega : E0.
Proof.  now exists omega%t1. 
Defined.

Notation omega  := _Omega.
\end{Coqsrc}

@@

\begin{Coqsrc}
Fixpoint succb (alpha:T1) : bool :=
  match alpha with
    | zero => false
    | ocons zero _ _ => true
    | ocons alpha n beta => succb beta
  end.

Fixpoint limitb (alpha : T1) : bool :=
  match alpha with
    | zero => false
    | ocons zero _ _ => false
    | ocons alpha n zero => true
    | ocons alpha n beta => limitb beta
  end.
\end{Coqsrc}

\begin{Coqsrc}
  Compute limitb omega.
\end{Coqsrc}

\begin{Coqsrc}
Compute succb 42.
\end{Coqsrc}

The correctness of these definitions with respect to the mathematical notions of
limit and successor ordinals is established through several lemmas. For instance,
Lemma \texttt{canonS\_limit}, page~\pageref{lemma:canonS-limit}, shows that
if $\alpha$ is (syntactically) a limit ordinal, then it is the least upper bound of
a strictly increasing sequence of ordinals.

   The following function is very useful in constructions by cases (proofs and function definitions).

\begin{Coqsrc}
Definition zero_succ_limit (alpha: T1) :
    {succb alpha} + {limitb alpha} +  {alpha=zero}.
    (* ... *)
Defined.
\end{Coqsrc}

@@

\begin{Coqsrc}
Fixpoint succ (alpha:T1) : T1 :=
  match alpha with 
   | zero => 1
   | ocons zero n _ => ocons zero (S n) zero
   | ocons beta n gamma => ocons beta n (succ gamma)
 end.
\end{Coqsrc}

@@

\index{hydras}{Exercises} 
\begin{exercise}
  Prove that that the ordinal addition on \texttt{E0} is associative.

  \textbf{Warning:}  Please do not import theorems from
  \texttt{ordinals/Schutte/}, since that library uses axioms!
\end{exercise}

@@

% \subsubsection{The total order \texttt{lt} on \texttt{T1} is \emph{not} well-founded}
% \label{sec:orgheadline76}

% Let us recall that the data type \texttt{T1} contains too many inhabitants, including
% terms which are not in Cantor normal form. Thus, the following result is not 
% very surprising.

% \begin{Coqsrc}
% Section lt_not_well_founded.

%   (* let us build the sequence of terms :
%         omega + omega + .... + omega ^ 2   *)
%   Let f := (fix f (i:nat): T1 :=
%             match i with 0 => phi0 2
%                        | S i => ocons 1 1 (f  i)
%             end).

%  Lemma  f_decreases : forall i, f (S i) <  f i.
%  Proof.
%   induction i; compute; auto with T1.
%  Qed.

%  Theorem lt_not_wf : ~  well_founded lt.
%  Proof. 
%    intro wf; case (not_decreasing _ lt);auto.
%    exists f; apply f_decreases.
%  Qed.

% End lt_not_well_founded.
% \end{Coqsrc}

% Thus, we have

@@

\begin{Coqbad}
Section First_attempt.

 Lemma wf_LT : forall alpha,  nf alpha -> Acc LT alpha. 
 Proof.
  induction alpha as [| beta IHbeta n gamma IHgamma].
  - split.
    inversion 1.
    destruct H2 as [H3 _];not_neg H3.
  -  split; intros delta Hdelta.
\end{Coqbad}

@@

\begin{Coqbad}
  Abort.
\end{Coqbad}

@@

>>>>

\index{Coq}{Unicity of equality proofs}

\label{sect:eq-proof-unicity}

@@

\input{movies/snippets/E0/E0Def}

@@

\input{movies/snippets/E0/E0Scope}

@@

\input{movies/snippets/E0/LtDef}

@@

\input{movies/snippets/E0/nfProofUnicity}

\input{movies/snippets/E0/E0EqIff}

@@

\label{sect:omega-T1}
\input{movies/snippets/E0/ZeroOmega}

%\index{Constants!zero:T1}

@@

\input{movies/snippets/T1/succbLimitb}

The correctness of these definitions with respect to the mathematical notions of
limit and successor ordinals is established through several lemmas. For instance,
Lemma \texttt{canonS\_limit\_lub}, page~\pageref{lemma:canonS-limit}, shows that
if $\alpha$ is (syntactically) a limit ordinal, then it is the least upper bound of
a strictly increasing sequence of ordinals.

@@

\begin{Coqsrc}
Fixpoint succ (alpha:T1) : T1 :=
  match alpha with 
   | zero => 1
   | ocons zero n _ => ocons zero (S n) zero
   | ocons beta n gamma => ocons beta n (succ gamma)
 end.
\end{Coqsrc}
%
% TO do : insert succ in E0
%

@@

@@

@@

\input{movies/snippets/T1/wfLTBada}

@@

@@

\input{movies/snippets/T1/wfLTBadz}
